---
title: Microeconometrics, Empirical project, Group 8
author:
- Atanasov Georgi^[student ID]
- Fitter Jonathan^[student ID]
- Geyer Niklas^[student ID]
- Hochholzer Matthias^[student ID  11724853]
- Woharcik Verena^[student ID]
date: 7th February 2021
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lastpage}
- \usepackage{graphicx}
- \pagestyle{fancy}
- \fancypagestyle{plain}{\pagestyle{fancy}}
- \fancyhead[L]{Empirical project} 
- \fancyhead[R]{\thepage\ of \pageref{LastPage}} 
- \fancyfoot[R]{\includegraphics[width=3cm]{Uni_Logo_blau.png}} 
- \fancyfoot[C]{}
- \setlength{\footskip}{46.27646pt}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
#Loading packages
library(tidyverse)
library(foreign)
library(ggplot2)
library(ROCR)
```


## Importing data
from Wooldridge, his source: J. Grogger (1991), “Certainty vs. Severity of Punishment,” Economic Inquiry
29, 297-309.
```{r}
df<-read.dta("http://fmwww.bc.edu/ec-p/data/wooldridge/crime1.dta")
attach(df)
head(df)
str(df)
summary(df)
```
A data.frame with 2725 observations on 16 variables:
- narr86: times arrested, 1986
- nfarr86: felony arrests, 1986
- nparr86: property crme arr., 1986
- pcnv: proportion of prior convictions
- avgsen: avg sentence length, mos.
- tottime: time in prison since 18 (mos.)
- ptime86: mos. in prison during 1986
- qemp86: quarters employed, 1986
- inc86: legal income, 1986, $100s
- durat: recent unemp duration
- black: =1 if black
- hispan: =1 if Hispanic
- born60: =1 if born in 1960
- pcnvsq: pcnv^2
- pt86sq: ptime86^2
- inc86sq: inc86^2

### Correlation Plots

```{r}
plot(df[,c("narr86", "avgsen", "inc86", "durat")])
cor(df[,c("narr86", "avgsen", "inc86", "durat")])
```

### Specific Plots:

```{r, echo=FALSE}
plot(narr86, inc86, main = "Correlation, crime 1986", xlab= "times arrested", ylab="legal income", col="darkblue")
```

```{r, echo=FALSE}
plot(narr86, avgsen, main = "Correlation, crime 1986", xlab= "times arrested", ylab="avg sentence length, mos.", col="darkblue")
```

```{r, echo=FALSE}
barplot(table(ifelse(df$black==1,df$narr86,NA)), xlab = "times arrested", ylab = "amount", main = "Histogram Black/White", col="blue")
barplot(table(ifelse(df$black==0,df$narr86,NA)), xlab = "times arrested", ylab = "amount", main = "Histogram Black/White", col="red")
```

## Simple OLS Regression, LPM

### OLS estimation

```{r, echo=FALSE}
# with all viariables
ols_all <- lm(narr86 ~ pcnv + avgsen + tottime + ptime86 + qemp86 + inc86 + durat + black + hispan + born60 + pcnvsq + pt86sq + inc86sq , data=df)
summary(ols_all)
```

### Model:
$$ narr86 = \beta_0 + \beta_1\ pcnv + \beta_2\ ptime86 +  \beta_3\ inc86 +  \beta_4\ black +  \beta_5\ hispan +  \beta_6\ pcnvsq+ \beta_7\ pt86sq+ \beta_8\ inc86sq   $$

```{r, echo=FALSE}
#with chosen variables
ols <- lm(narr86 ~ pcnv + ptime86 + inc86 + black + hispan + pcnvsq + pt86sq + inc86sq  , data=df)
summary(ols)
```

## IV Regression

### Model:



### Generalized IV estimation


### 2SLS estimation


## LOGIT model
creating a binary variable arr86, when a person gets arrested at least once.
```{r}
df$arr86 <- ifelse(df$narr86>0 ,1 ,0)
```

```{r}
log_all <- glm(arr86 ~ pcnv + avgsen + tottime + ptime86 + qemp86 + inc86 + durat + black + hispan + born60 + pcnvsq + pt86sq + inc86sq , data = df, family=binomial(link = "logit"))

summary(log_all)
```
```{r}
log <- glm(arr86 ~   ptime86 + qemp86 + inc86  + black + hispan + pcnvsq + pt86sq + inc86sq , data = df, family=binomial(link = "logit"))

summary(log)
```

For comparison
```{r}
prob <- glm(arr86 ~   ptime86 + qemp86 + inc86  + black + hispan + pcnvsq + pt86sq + inc86sq , data = df, family=binomial(link = "probit"))

summary(prob)
```

## Calculation of MC Faddens pseudo R^2
```{r}
1-(prob$deviance/prob$null.deviance)
```



## parameter logit probit check
prob/log
slide 21 ... factor 1.6

##Average marginal effect
```{r}
# for logit
fav <- mean(dnorm(predict(log,type="link")))
fav*coef(log)
```

## Predicition of cutoff
```{r}

```


## ROC
```{r}
pred <- prediction(fitted(log),df$arr86)
plot(performance(pred, "acc"))

plot(performance(pred,"tpr", "fpr"))
abline(0,1,lty=2)
```



## Truncated model, let's see
